---
title: "MATH1324 Assignment 1"
subtitle: Statistical analysis of Climate data
output:
  pdf_document: default
  html_notebook: default
---

## Student Details

Wing Hang Chan (s3939713)

Matthew Ng Jun Hung (s3963313)

Sukhum Boondecharak (s3940976)

## Problem Statement

Many people in Melbourne and Sydney suffered heatstroke in last summer. For example, Rafael Nadal reportedly suffered heat stroke and lost 4 kilograms in Australian Open (Ref 1). Heatstroke (Ref 2) is a condition caused by your body overheating, usually as a result of prolonged exposure to or physical exertion in high temperatures. It is the main reason we have chosen max temperature and solar exposure as variables. High temperatures and high solar exposure may relatively be the cause of body overheating.

Ref 1: Rafael Nadal reportedly suffered heat stroke - <https://7news.com.au/sport/tennis/rafael-nadal-reportedly-suffered-heat-stroke-lost-4kg-in-australian-open-win-over-shapovalov-c-5457236>

Ref 2: Heatstroke condition - <https://www.mayoclinic.org/diseases-conditions/heat-stroke/symptoms-causes/syc-20353581>

## Load Packages

```{r}
library(dplyr)
```

## Data

```{r}
data_mel <- read.csv('Climate Data-Melbourne-1.csv')
data_syd <- read.csv('Climate Data-Sydney-1.csv')

colnames(data_mel)[colnames(data_mel) == 'Maximum.temperature..Degree.C.'] <- 'max_temp'
colnames(data_mel)[colnames(data_mel) == 'solar.exposure'] <- 'solar_exp'

colnames(data_syd)[colnames(data_syd) == 'Maximum.temperature..Degree.C.'] <- 'max_temp'
colnames(data_syd)[colnames(data_syd) == 'solar.exposure'] <- 'solar_exp'

min_temp <- min(min(data_mel$max_temp), min(data_syd$max_temp))
max_temp <- max(max(data_mel$max_temp), max(data_syd$max_temp))

x_lim_max_temp = c(min_temp, max_temp)

min_solar_exp <- min(min(data_mel$solar_exp), min(data_syd$solar_exp))
max_solar_exp <- max(max(data_mel$solar_exp), max(data_syd$solar_exp))

x_lim_solar_exp = c(min_solar_exp, max_solar_exp)
```

## Summary Statistics

Calculate mean, median, standard deviation, first and third quartile, interquartile range, minimum and maximum values) of the selected variable grouped by city.

```{r}
#Statistic for Maximum Temperature in Melbourne

data_mel %>% summarise(Mean = mean(max_temp), Median = median(max_temp),
                       Stand_dev = sd(max_temp),
                       First_quarter = quantile(max_temp, 0.25),
                       Third_quarter = quantile(max_temp,0.75),
                       Interquartile = IQR(max_temp),
                       Min = min(max_temp), Max = max(max_temp)) 

#Statistic for Solar Exposure in Melbourne

data_mel %>% summarise(Mean = mean(solar_exp), Median = median(solar_exp),
                       Stand_dev = sd(solar_exp),
                       First_quarter = quantile(solar_exp, 0.25),
                       Third_quarter = quantile(solar_exp,0.75),
                       Interquartile = IQR(solar_exp),
                       Min = min(solar_exp), Max = max(solar_exp)) 

#Statistic for Maximum Temperature in Sydney

data_syd %>% summarise(Mean = mean(max_temp), Median = median(max_temp),
                       Stand_dev = sd(max_temp),
                       First_quarter = quantile(max_temp, 0.25),
                       Third_quarter = quantile(max_temp,0.75),
                       Interquartile = IQR(max_temp),
                       Min = min(max_temp), Max = max(max_temp)) 

#Statistic for Solar Exposure in Sydney

data_syd %>% summarise(Mean = mean(solar_exp), Median = median(solar_exp),
                       Stand_dev = sd(solar_exp),
                       First_quarter = quantile(solar_exp, 0.25),
                       Third_quarter = quantile(solar_exp,0.75),
                       Interquartile = IQR(solar_exp),
                       Min = min(solar_exp), Max = max(solar_exp)) 

```

## Distribution Fitting

```{r}
#Distribution of variables
#Histogram of Melbourne's Max Temperature
# par(mfrow = c(1, 1))
#par(mfrow = c(2, 2))
hist(data_mel$max_temp,
     breaks = 30,
     main = "1: Histogram of Melbourne's Max Temperature", 
     xlab = "Max Temperature in Celsius"  ,
     xlim = x_lim_max_temp,
     prob = T,
     las = 1) 
lines(density(data_mel$max_temp, adjust = 2), col = "Red", lwd = 2)
abline(v=mean(data_mel$max_temp),col='black',lw=2)

#Normal Distribution Data Fitting of Melbourne's Max Temperature
#Blue Line
lines(seq(0,
          max(data_mel$max_temp), by=.5), 
      dnorm(seq(0,
                max(data_mel$max_temp), by=.5), 
            mean(data_mel$max_temp), 
            sd(data_mel$max_temp)), col="Blue")

legend(x = 'topright',
       legend = c("Empirical Density","Normal Density","Mean"),
       col = c('red',"Blue","black"),
       lty = c(1,1,1))

#Histogram of Melbourne's Solar Exposure
hist(data_mel$solar_exp,
     breaks = 30,
     main = "2: Histogram of Melbourne's Solar Exposure",
     xlab = "Solar Exposure",
     xlim = x_lim_solar_exp,
     prob = T,
     las = 1)
lines(density(data_mel$solar_exp, adjust = 2), col = "Red", lwd = 2)
abline(v=mean(data_mel$solar_exp),col='black',lw=2)

#Normal Distribution Data Fitting of Melbourne's Solar Exposure
lines(seq(min(data_mel$solar_exp),
          max(data_mel$solar_exp), by=.5), 
      dnorm(seq(min(data_mel$solar_exp),
                max(data_mel$solar_exp), by=.5), 
            mean(data_mel$solar_exp), 
            sd(data_mel$solar_exp)), col="Blue")

legend(x = 'topleft',
       legend = c("Empirical Density","Normal Density","Mean"),
       col = c('red',"Blue","Black"),
       lty = c(1,1,1),
       cex = 0.5)

#Histogram of Sydney's Max Temperature
hist(data_syd$max_temp,
     breaks = 30,
     main = "3: Histogram of Sydney's Max Temperature",
     xlab = "Max Temperature in Celsius",
     xlim = x_lim_max_temp,
     prob = T,
     las = 1)
lines(density(data_syd$max_temp, adjust = 2), col = "Red", lwd = 2)
abline(v=mean(data_syd$max_temp),col='black',lw=2)

#Normal Distribution Data Fitting of Sydney's Maximum Temperature
lines(seq(0,
          max(data_syd$max_temp), by=.5), 
      dnorm(seq(0,
                max(data_syd$max_temp), by=.5), 
            mean(data_syd$max_temp), 
            sd(data_syd$max_temp)), col="Blue")

legend(x = 'topright',
       legend = c("Empirical Density","Normal Density","Mean"),
       col = c('red',"Blue","Black"),
       lty = c(1,1,1),
       cex = 0.7)

#Histogram of Sydney's Solar Exposure
hist(data_syd$solar_exp,
     breaks = 30,
     main = "4: Histogram of Sydney's Solar Exposure",
     xlab = "Solar Exposure",
     xlim = x_lim_solar_exp,
     prob = T,
     las = 1)
lines(density(data_syd$solar_exp, adjust = 2), col = "Red", lwd = 2)
abline(v=mean(data_syd$solar_exp),col='black',lw=2)

#Normal Distribution Data Fitting of Sydney's Solar Exposure
lines(seq(0,
          max(data_syd$solar_exp), by=.5), 
      dnorm(seq(0,
                max(data_syd$solar_exp), by=.5), 
            mean(data_syd$solar_exp), 
            sd(data_syd$solar_exp)), col="Blue")


legend(x = 'topleft',
       legend = c("Empirical Density","Normal Density","Mean"),
       col = c('red',"Blue","black"),
       lty = c(1,1,1),
       cex = 0.7) 
```

## Interpretation

As graph 1 and 3 show a flatter curve compared with normal distribution curve, it may be because we got a small size of data set. There is no evidence that maximum temperature in both countries relates to a high chance of people suffering from heatstroke with the balance distribution. However, there is a finding with analysing data of solar exposure. It shows both cities had much sunshine last summer as the mean value is at the right hand side of the graph. There is a higher chance to have higher solar exposure than normal expected in both Melbourne and Sydney. It may explain why many people suffered from heatstroke last summer.
